---
title: 'M3: R for Data Science Chapter 3'
output:
  html_document: default
date: '2022-07-03'
---

This text will show up when this is "knitted" (if you look above in RStudio, you will see a "Knit" button).

# Setup

This "chunk" will load any necessary packages and set any run-time options for this document. 

```{r setup, include=FALSE}
# load the libraries within the tidyverse 
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE)
```

When `ggplot` is called, it will return an object that can be passed to further functions.

But if you just call `ggplot` without assigning it, this notebook will assume you want
to display it. Which means if you run this next chunk it will show the visualizations 
in line in the notebook.

The cars which get less gas mileage on the highway have more engine displacement. Vehicles that have an engine with higher displacement can create more power which leads to more fuel consumption. On the other hand engines with lower displacement consume less fuel but aren't able to create as much power. This information would prove useful to strategic groups within the automobile industry when trying to evaluate new opportunities for financial and competitive growth. 

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))
```

# 3.2.4 Exercises

From [Exercises in Section 3.2.4](https://r4ds.had.co.nz/data-visualisation.html#exercises).

## Run `ggplot(data = mpg)`. What do you see and why?
After running ggplot(data = mpg) a empty graph was created. This is because in order for the graph to be completed a geom and mapping must be specified. Something such as geom_point(mapping = aes(x = displ, y= hwy)). Where geom_point adds the layer of points to the plot, and the mapping argument determines how the variables from the dataset are displayed. As stated in the textbook the "mapping argument is always paired with aes()."The x and y arguments of aes() are what specify which variables to put on the x and y axes. 

```{r}
ggplot(data = mpg)
```

## How many rows are in mpg? How many columns?
There are 234 rows and 11 columns.

Hint: check out the help for`nrow()` and `ncol()` and the `dim()` functions.

```{r}
nrow(mpg)
ncol(mpg)
dim(mpg)
```


## What does the drv variable describe? Read the help for ?mpg to find out.

Well the drv variable represents the type of drive train, where f indicates the vehicle is front-wheel drive, r = rear wheel drive and 4 = four-wheel drive. The drive train variable is a variable that's specifications have an important impact on a vehicles fuel economy.  

## Make a scatterplot of `hwy` vs `cyl`.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = cyl, y= hwy))
```

## What happens if you make a scatterplot of class vs drv? Why is the plot not useful?
When making a scatterplot of the class and drv, the only information that would be displayed is the combination of class type and drive type such as a front wheel drive, midsize vehicle.

```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = drv, y= class))
```


# 3.2.4 Exercises

## What’s gone wrong with this code? Why are the points not blue?
The "color = 'blue'" needs to be placed outside of the aes() argument.

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = "blue"))
#After placing color = "blue" outside of the aes() argument, you can see that each of the points on the scatterplot are now blue. 
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

## Which variables in mpg are categorical? Which variables are continuous? (Hint: type ?mpg to read the documentation for the dataset). How can you see this information when you run mpg?
The character strings such as manufacturer, model, drv, trans, fl and class are categorical variables. One way to view this information is by running the code 'str(mpg)' which shows the variable types (int, chr, num.  

```{r}
?mpg
str(mpg)
```


## Map a continuous variable to color, size, and shape. How do these aesthetics behave differently for categorical vs. continuous variables?
These aesthetics behave differently as when plotting a continuous variables such as displ to color, shape, or size.ggplot creates a color scale in which this color scale represents the values of the continuous variable. 


```{r}
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, color = displ))
ggplot(data = mpg) +
  geom_point(mapping = aes(x = displ, y = hwy, size = displ))
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy)) + 
  facet_wrap(~ class, nrow = 2)
```


## What happens if you map the same variable to multiple aesthetics?
More then one aesthetic will be implemented allowing for multiple legends to be generated. 

(*Hint: Run the code below*)

```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = drv, shape=drv))
```


## Describe what's happening with the color aesthetic in the following code.
displ < 5 is being evaluated. For any number less than 5 a blue dot is given which represents true and for any value greater than 5 a red dot is shown which represents false. displ < 5 creates a temporary boolean variable which indicates whether the observations satisfy the set conditions. 
```{r}
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy, color = displ < 5))
```

# 3.6.1 Exercises

## What geom would you use to draw a line chart? A boxplot? A histogram? An area chart?

 * Line Chart: geom_line()
 * Box plot: geom_boxplot()
 * Histogram: geom_histogram()
 * Area Chart: geom_area() 


## prediction vs reality

Run this code in your head and predict what the output will look like. Then, run the code in R and check your predictions.


```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point() + 
  geom_smooth(se = FALSE)

```

## What does show.legend = FALSE do? What happens if you remove it?
It hides the legend provided by default and if we were to remove it the result would be he legend being displayed in the graph. 

```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point(show.legend = FALSE) + 
  geom_smooth(method="loess", formula = y~x, se = FALSE, show.legend = FALSE)

ggplot(data = mpg, mapping = aes(x = displ, y = hwy, color = drv)) + 
  geom_point(show.legend = TRUE) + 
  geom_smooth(method="loess", formula = y~x, se = TRUE, show.legend = TRUE)

```


## Before Running: do you think these two graphs look different? Why/why not?
Before running the code I believed that the graphs would look the same. The only difference between the two codes is that the first set of code defines data and mapping in ggplot. In the second graph data and mapping are defined separately
in each geom. 
```{r}
ggplot(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_point() + 
  geom_smooth()

ggplot() + 
  geom_point(data = mpg, mapping = aes(x = displ, y = hwy)) + 
  geom_smooth(data = mpg, mapping = aes(x = displ, y = hwy))
```


# 3.7.1 Exercises

## What does geom_col() do? How is it different to geom_bar()?
geom_col and geom_bar both create bar charts. geom_col uses the default geom_identity where the heights of the bars represent the values within the data. geom_bar uses stat_count which counts the frequencies within each grou and adjusting the heights of the bars equal to those frequeincies. ?

## What variables does stat_smooth() compute? What parameters control its behaviour?

stat_smooth computes the variables y,x (predicted value), ymin/xmin which is the lower pointwise confidence intervale around the mean. It also computes ymax/xmax which is the upper pointwise confidence interval around the mean and the standard error (se). 

The parameters that control the behavior of stat_smooth are the n, span, fullrange, and level controls. n specifies the number of points at which to evaluate smoother, span- controls the amount of smoothing for the default loess smoother, fullrange- should the fit span the full range of the plot or just the data?, and level- which is the level of confidence interval to use which is usually set to 0.95 by default. 


# 3.8.1 Exercises

## What is the problem with this plot? How could you improve it?
The issue is that the values of cty and hwy are rounded so the points appear on the grid but many of them overlap which is an issue known as overplotting. One way to improve this is through setting the postion adjustment to jitter. This adds a small amount of random noise to each point, thus spreading them out as no two points will have the same amount of random noise. Through adding this randomness the graph will be less accurate at small scales but makes the graph more revealing at large scales. 

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + 
  geom_point()
```

## What parameters to geom_jitter() control the amount (up/down and left/right) of jittering?
The width and height arguments can be used to control the amount of jittering both vertically and horizontally in both positive and negative directions. 


# 3.9.1 Exercises

## Turn the following stacked bar chart into a pie chart using coord_polar().

```{r}
ggplot(mpg, aes(x=class, fill=drv)) + 
  geom_bar(width = 1) +
  coord_polar(theta = "y")

bar <- ggplot(data = diamonds) + 
  geom_bar(
    mapping = aes(x = cut, fill = cut), 
    show.legend = FALSE,
    width = 1
  ) + 
  theme(aspect.ratio = 1) +
  labs(x = NULL, y = NULL)
bar + coord_flip()
bar + coord_polar()# uses polar coordinates which reveal connection between barchart and Coxcomb chart 

```


## Describe what the ggplot function `labs()` does when plotting? 
The ggplot function 'labs()' allows us to modify/change the axis, legend as well as plot labels within a graph. It is important to make sure the axis and legend labels display the full variable name. Good labels are very important when it comes to making your plots available to a wider audience. 

(Read the help page and use your own words)
